---
output:  
  html_document:  
    keep_md: true  
---

# Basic R Exercises

#### Create the following dataframe

transactionID | Marketing Channel | Product | Returns | Sales
--------------|-------------------|---------|--------|------
1001 | Amazon | Macbook | Yes | 1200
1002 | Amazon | Macbook | Yes | 1200
1003 | Amazon | Macbook | Yes | 1200
1004 | Amazon | Macbook | No | 1200
1005 | Amazon | Macbook | No | 1200
1006 | Best Buy | Macbook | Yes  | 1000
1007 | Best Buy | Macbook | No | 1000
1008 | Best Buy | Macbook | No | 1000
1009 | Best Buy | Macbook | No | 1000
1010 | Best Buy | Macbook | No | 1000

```{r}

transactionID <- seq(from=1001,to=1010,by=1)
marketingChannel <- c(rep("Amazon",5),rep("Best Buy",5))
product <- rep("Macbook",10)
returns <- c("Yes","Yes","Yes","No","No","Yes","No","No","No","No")
sales <- c(rep(1200,5),rep(1000,5))

(salesData <- data.frame(transactionID,marketingChannel,product,returns,sales))

```

* Change the variables to have the following data types
    + transactionID= factor
    + Marketing Channel= Factor
    + Product= Factor
    + Returns= Logical
    + Sales= numeric
Hint: Yes and No in returns must be transformed to TRUE FALSE
    
```{r}

salesData$transactionID <- factor(salesData$transactionID)
salesData$returns <- ifelse(salesData$returns=="Yes",TRUE,FALSE)
salesData$returns <- as.logical(salesData$returns)

str(salesData)

```


* Change four of the Sales figues to NA without creating a new dataframe

```{r}
salesData[c(1,3,5,7),5] <- NA
```

* Using is.na, change the NA values to be the mean of the sales column

```{r}
salesData[is.na(salesData[,5]),5] <- mean(salesData$sales,na.rm=TRUE)
```

* Reduce the sales value for all returned items by 30%

```{r}
salesData$sales <- ifelse(salesData$returns==TRUE,0.7*salesData$sales,salesData$sales)
```

* Return a table that shows the total sales by marketingChannel

```{r}
(totalAmazon <- sum(salesData[1:5,5]))
(totalBestBuy <- sum(salesData[6:10,5]))
```
