---
output:  
  html_document:  
    keep_md: true  
---

# Exploratory Analysis Exercises   

## Simulating Dice Rolls

  * Simulate a 2 die roll for 100 trials
  * Plot a histogram of the results
    + Use relative frequency instead of count
    + Add color and formatting to the histogram plot
    + Add an axis to include all values between 2:12


In order to create a simulation of dice rolls we use sample() to take a random sample of integers between 1 and 6.  If an analysis were to be run at different times or by various individuals it could be extremely tedious if the results were always different.  Although replication is not an issue with this exercise, we use set.seed so that the analysis will always produce the same results.
```{r}
set.seed(98250)
```

Create a dataframe of NA's that will house the dice trials.  Each trial will have a value for roll 1, roll 2, and a total of both rolls.  Thus, we need 3 columns and we'll run 100 trials. 
```{r}
rolls <- data.frame(matrix(NA,ncol=3,nrow=100))   
names(rolls) <- c("roll1","roll2","Total")
```

For rolls 1 and 2, use sample() to pick a number between 1 and 6 for 100 trials.  Every time we roll the die all 6 sides are available which means that we need to set replace to true.  Otherwise, if we roll a three on the first trial it would not be included in the sample set for the following rolls.
```{r}
rolls$roll1 <- sample(x=c(1:6) , size=100 , replace=TRUE)
rolls$roll2 <- sample(x=c(1:6) , size=100 , replace=TRUE)
rolls$Total <- rolls$roll1+rolls$roll2

rolls$Total <- as.numeric(rolls$Total)
```


To visualize the data we create a histogram.  The bins are expliclity states using breaks=.
```{r}
hist(rolls$Total,
#      breaks=10,
     breaks=c(seq(2,12,1)),
     main="Simulation of Two Dice Rolls",
     xlab="Sum of two rolls",
     ylab="Relative Frequency",
     col="lightblue",
     xaxt="n",
     freq=FALSE
)
axis(side=1, at=seq(2,12,by=1))
```




## Analyzing Philly Dropout Data available from 

Data obtained from https://www.opendataphilly.org/dataset/schools

* Load the file Philly_Dropout_Data.csv
* Remove the following columns: (2,3,4,6)
* Name the columns:
    + County FACTOR
    + School CHARACTER
    + Enrollments INTEGER
    + maleDropouts INTEGER
    + femaleDropouts INTEGER
    + dropoutRate NUMERIC
* Data from opendataphilly.org

```{r}

data <- read.csv("http://goo.gl/FieXmC")
data <- data[,c(-2,-3,-4,-6)]
names(data) <- c("county","school","enrollments","maleDropouts","femaleDropouts","dropouts","dropoutRate")

data$county <- as.factor(data$county)
data$school <- as.character(data$school)

str(data)

```

  * Remove missing values from the dataset

```{r]}
#data <- data[!is.na(data),]
data <- na.omit(data)
summary(data)
```

* Show descriptive statistitcs for the dataset.
* Create a table that aggregates enrollments, dropouts,maleDropouts,femaleDropouts by county

```{r}

# describe(data[,c(1,4:7)])

aggData <- data.frame(aggregate(cbind(enrollments,dropouts,maleDropouts,femaleDropouts)~county,data=data,sum))
str(aggData)
head(aggData);tail(aggData)

```

* Compute the average, min, max dropout rate per county omitting schools with 0 dropout rates
```{r}

newData <- data[data$dropouts>0,]
newData <- data[data$dropouts<100,]


countyMed <- aggregate(dropouts~county,data=newData,mean)
countyMin <- aggregate(dropouts~county,data=newData,min)
countyMax <- aggregate(dropouts~county,data=newData,max)

(dropData <- cbind(countyMin,countyMed$dropouts,countyMax$dropouts))

```

* Create a boxplot comparing male dropouts and female dropouts using the new dataset
```{r}
newData <- data[data$dropouts>5&data$dropouts<100,]

boxplot(cbind(newData$maleDropouts,newData$femaleDropouts) , xaxt="n", 
        ylab="Number of Dropouts",
        main="Number of Dropouts by Gender")
axis(side=1, at=c(1,2) , labels=c("Male","Female"))

```


